The 2018 heatwave and its implications on ozone induced damage on vegetation in a subarctic climate

### 2020-03-26 VegOzone.tex
Some more text in the data section and a new plot dedicaded to monitoring data from Svanvik in 2018/19.

In our assessment on ozone damage on vegetation is the subarctic in relation to the 2018 heatwave, we focus on the NIBIO operated local site Svanhovd which lies close by the settlement of Svanvik in the Pasvik valley. Throughout the paper, we will use Svanvik synonymously for the atmospheric monitoring site at Svanhovd. For the purpose of this study, we evaluate data taken at Svanhovd (Section~\ref{subsec:svanhovd}) with observational data from other monitoring sites in northern Fennoscandia (Section~\ref{subsec:obs_data}). Reanalysis data is used as complement where observational data was not available or temporally sparse (Section~\ref{subsec:model_data}).

\subsection{NIBIO research station Svanhovd}
\label{subsec:svanhovd}
"Svanhovd is located beside the Pasvik river, in the middle of the wedge of Norwegian land separating Russia and Finland in the north. It is where the eastern Siberian taiga meets the western Boreal forest. [...] Svanhovd was originally established as a State Demonstration and Experimental Farm in the 1930's, in order to expand agriculture in Pasvik valley [...]" - \href{https://www.nibio.no/en/about-eng/addresses/northern-norway/svanhovd}{NIBIO}\\ The research station and locations of the atmospheric monitoring site and the ozone garden are shown in the aerial photography (Fig.~\ref{fig:station_map_fennoscandia}). Svanhovd comprises atmospheric monitoring (ozone monitoring exclusively for the summer seasons 2018/19) and observed ozone damage in plants.
\emph{Comment: Invite co-authors from NILU and NIBIO?}

@@

\subsubsection{Ozone monitoring}
\label{subsubsec}
{\bf TODO: (Technical) details about ozone monitoring at Svanhovd - historical period. Project related monitoring (summer seasons 2018/19.) $\rightarrow$ NILU -- Sverre?}

@@

\subsubsection{Surface ozone observations}
\label{subsubsec:ebas}
All surface ozone observations used in this study are available from NILU EBAS database except for the data for the 2018/19 summer season at Svanhovd. In Table~\ref{tab:ebas_obs}, we give an overview over the exact locations and time range of operational data taking.

\begin{table}[t]
  \caption{Past and present ozone observation sites in northern Fennoscandia. Data available on \href{http://ebas.nilu.no/}{EBAS}.}
  \label{tab:ebas_obs}
  \begin{tabular}{lllllrl}
    \tophline
    Name       & Country & ID      & \multicolumn{3}{c}{Location} & Operational\\
    &         &         & lat             & lon               & alt            &\\
    &         &         & (\unit{^\circ N}) & (\unit{^\circ E})  & (\unit{m})     &\\
    \middlehline
    Esrange    & SWE     & SE0013R & 67.83           & 21.07             & 475            & 1991 -- present\\
    Janiskoski & RUS     & RU0001R & 68.93           & 28.85             & 118            & 1995 -- 1997\\
    Jergul     & NOR     & NO0030R & 69.45           & 24.60             & 255            & 1997 -- 2011\\
    Karasjok   & NOR     & NO0055R & 69.467          & 25.217            & 333            & 1988 -- 1997\\
    Pallas     & FIN     & FI0096G & 69.97           & 24.12             & 565            & 1995 -- present\\
    Svanvik    & NOR     & NO0047R & 69.45           & 30.03             & 30             & 1986 -- 1996$^\dagger$\\
    \bottomhline
  \end{tabular}
  \belowtable{$^\dagger$ Exclusive ozone monitoring summer seasons 2018/19 operated by NILU.} % Table Footnotes
\end{table}

\begin{figure*}[t]
  \includegraphics[width=12cm]{ozone_timeseries_fenoscandic_obs.png}
  \label{fig:ozone_timesseries_fenoscandic_obs}
  \caption{Time series of ozone observations in northern Fennoscandia. Data taken from EBAS as well as from dedicated ozone monitoring at Svanhovd in the summer seasons 2018/19 (Tab.~\ref{tab:ebas_obs}). The hatched areas indicate time periods with insufficient quality control - \bf{TODO: citation NILU ozone report 200x}.}
\end{figure*}

\subsubsection{Atmospheric monitoring}
Monitoring data from Svanvik/Pasvik.
Ozone/Temperature (2018/2019): Luftkvalitet.no (NILU), LandbruksMeteorologiskTjeneste (NIBIO)
We refrain from doing a full cross-evaluation for these meteorological fields, since this has been done elsewhere {\bf TODO: It's somewhat clear that this is a job of the operators in the monitoring network. Find a citation for this.}.
\begin{itemize}
\item Temperature
\item Precipitation
\item Incoming radiation \emph{Comment to Ane: Same plot that you've included in the Nordic ozone report?}
\item ...
\end{itemize}

>>>>

In our assessment on ozone damage on vegetation in the subarctic in relation to the 2018 heatwave, we focus on the NIBIO operated local site Svanhovd which lies close by the settlement of Svanvik in the Pasvik valley (Section~\ref{subsec:svanhovd}). Throughout the paper, we will use Svanvik synonymously for the atmospheric monitoring site at Svanhovd. The available agrameteorological monitoring data are described in Section~\ref{subsec:atmo_svanvik}. In Section~\ref{subsubsec:ozone_garden}, the ozone garden -- a plant-based ozone monitoring -- will be described in more detail. For evaluation of the ozone data taken at Svanhovd, we compare with observational data from other monitoring sites in northern Fennoscandia (Section~\ref{subsec:obs_data}). Reanalysis data is used as complement where observational data was not available or temporally sparse. The chosen data sets are described in Section~\ref{subsec:model_data}.

\subsection{NIBIO research station Svanhovd}
\label{subsec:svanhovd}
"Svanhovd is located beside the Pasvik river, in the middle of the wedge of Norwegian land separating Russia and Finland in the north. It is where the eastern Siberian taiga meets the western Boreal forest. [...] Svanhovd was originally established as a State Demonstration and Experimental Farm in the 1930's, in order to expand agriculture in Pasvik valley [...]" - \href{https://www.nibio.no/en/about-eng/addresses/northern-norway/svanhovd}{NIBIO}.\\
The research station at Svanhovd comprises long-term observations of various agrameteorological variables, i.a. $2\,\unit{m}$ temperature, precipitation, global radiation, and soil moisture and temperature at different depths. For the purpose of relating observed damage on ozone sensitive plants (e.g. clover and tobacco) an ozone monitor had been installed during the growing seasons 2018/19). The relevant data are described in more detail Section~\ref{subsubsec:atmo_svanvik}.
Note that from 1986--1996, ozone monitoring has been conducted operationally at Svanhovd. Details about these historical ozone data will be given in Section~\ref{subsubsec:ebas}.\\
The locations of the atmospheric monitoring site as well as the ozone garden with respect to the main buildings of the NIBIO research station are marked in the aerial photography shown in Fig.~\ref{fig:svanhovd_research_station}. 
\emph{Comment: Invite co-authors from NILU and NIBIO?}

@@

\subsubsection{Atmospheric monitoring}
\label{subsubsec:atmo_svanvik}
Monitoring data of key variables in the growing seasons 2018/19 have been downloaded from \href{luftkvalitet.no} operated by NILU and are shown in Fig.~\ref{fig:data_svanvik_2018_2019}. The hatched areas mark times when no ozone data has been taken. Note, while the downtime during winter was planed, missing data in two weeks of July 2018 were due to problems in data acquisition. For longer time series: LandbruksMeteorologiskTjeneste (NIBIO)

      {\bf TODO: (Technical) details about ozone monitoring at Svanhovd - historical period. Project related monitoring (growing seasons 2018/19.) $\rightarrow$ NILU -- Sverre?}

\emph{Comment: We refrain from doing a full cross-evaluation for these meteorological fields, since this has been done elsewhere. It's somewhat clear that this is a job of the operators in the monitoring network.} {\bf TODO: Find a citation for this.}.
\begin{itemize}
\item Incoming radiation \emph{Comment to Ane: Same plot that you've included in the Nordic ozone report?}
\item ...
\end{itemize}

\begin{figure*}[t]
  \includegraphics[width=12cm]{data_svanvik_2018_2019}
  \caption{Observational data from atmospheric monitoring at Svanhovd in 2018/19. The hatched areas indicate missing ozone monitoring data. (a) Ozone; (b) temperature; (c) daily accumulated precipitation.}
  \label{fig:data_svanvik_2018_2019}
\end{figure*}

@@

\subsubsection{Surface ozone observations}
\label{subsubsec:ebas}
All surface ozone observations used in this study are available from NILU EBAS database except for the data for the 2018/19 growing season at Svanhovd. In Table~\ref{tab:ebas_obs}, we give an overview over the exact locations and time range of operational data taking.

\begin{table}[t]
  \caption{Past and present ozone observation sites in northern Fennoscandia. Data available on \href{http://ebas.nilu.no/}{EBAS}.}
  \label{tab:ebas_obs}
  \begin{tabular}{lllllrl}
    \tophline
    Name       & Country & ID      & \multicolumn{3}{c}{Location} & Operational\\
    &         &         & lat             & lon               & alt            &\\
    &         &         & (\unit{^\circ N}) & (\unit{^\circ E})  & (\unit{m})     &\\
    \middlehline
    Esrange    & SWE     & SE0013R & 67.83           & 21.07             & 475            & 1991 -- present\\
    Janiskoski & RUS     & RU0001R & 68.93           & 28.85             & 118            & 1995 -- 1997\\
    Jergul     & NOR     & NO0030R & 69.45           & 24.60             & 255            & 1997 -- 2011\\
    Karasjok   & NOR     & NO0055R & 69.467          & 25.217            & 333            & 1988 -- 1997\\
    Pallas     & FIN     & FI0096G & 69.97           & 24.12             & 565            & 1995 -- present\\
    Svanvik    & NOR     & NO0047R & 69.45           & 30.03             & 30             & 1986 -- 1996$^\dagger$\\
    \bottomhline
  \end{tabular}
  \belowtable{$^\dagger$ Exclusive ozone monitoring growing seasons 2018/19 operated by NILU.} % Table Footnotes
\end{table}

\begin{figure*}[t]
  \includegraphics[width=12cm]{ozone_timeseries_fenoscandic_obs.png}
  \label{fig:ozone_timesseries_fenoscandic_obs}
  \caption{Time series of ozone observations in northern Fennoscandia. Data taken from EBAS as well as from dedicated ozone monitoring at Svanhovd in the growing seasons 2018/19 (Tab.~\ref{tab:ebas_obs}). The hatched areas indicate time periods with insufficient quality control - \bf{TODO: citation NILU ozone report 200x}.}
\end{figure*}
